---
title: "All PCA"
author: "Kyle Sullivan"
date: "4/16/2021"
output: html_document
---


```{r}
library(tidyverse)
library(reshape2)
library(factoextra)
library(FactoMineR)
library(adegenet)
```
make the total dataframe for ORF PCA
```{r}
total_pca_dataframe_ORF <- codon_frequency_table
```
Add the starting locations for each of the open reading frames
```{r}
SEE_starts = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SEE Genes/SEE_starts.txt", header = FALSE, sep= ",", check.names = FALSE) 

SR_starts = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SR Genes/SR_starts.txt", header = FALSE, sep= ",", check.names = FALSE) 

STC_starts = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/STC Genes/STC_starts.txt", header = FALSE, sep= ",", check.names = FALSE) 

STE_starts = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/STE Genes/STE_starts.txt", header = FALSE, sep= ",", check.names = FALSE) 

SEC_starts = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SEC Genes/SEC_starts.txt", header = FALSE, sep= ",", check.names = FALSE) 

locations <- c(SEE_starts, SR_starts, STC_starts, STE_starts, SEC_starts)

total_pca_dataframe_ORF$Location <- locations
```
Distiguish whether or not the open reading frame is located in a pathogenicity island
```{r}
total_pca_dataframe_ORF <- total_pca_dataframe_ORF %>% mutate(is_it_SPI1 = 
                                case_when(total_pca_dataframe_ORF$Location>=3005787 & total_pca_dataframe_ORF$Location<=3050270 ~ "SPI1",
                                          total_pca_dataframe_ORF$Location>=1461452 & total_pca_dataframe_ORF$Location<=1501953 ~ "SPI2",
                                          total_pca_dataframe_ORF$Location<3005787 & total_pca_dataframe_ORF$Location>1501953~ "other",
                                          total_pca_dataframe_ORF$Location > 3050270 ~ "other",
                                          total_pca_dataframe_ORF$Location < 1461452 ~ "other"))
```
Read in all of the GC contents for each open reading frame 
```{r}
SEE_GC = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SEE_GC_content.txt", header = FALSE, sep= ",", check.names = FALSE) 

SR_GC = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SR_GC_content.txt", header = FALSE, sep= ",", check.names = FALSE) 

STC_GC = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/STC_GC_content.txt", header = FALSE, sep= ",", check.names = FALSE) 

STE_GC = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/STE_GC_content.txt", header = FALSE, sep= ",", check.names = FALSE) 

SEC_GC = read.delim("~/Documents/University of Maryland/IVSP/Capstone/Genomic Data/Genomes/cds files/SEC_GC_content.txt", header = FALSE, sep= ",", check.names = FALSE) 

GC_contents = c(SEE_GC,SR_GC,STC_GC,STE_GC,SEC_GC)

total_pca_dataframe_ORF$GC_content = GC_contents
```
Add strain labels to each open reading frame
```{r}
SEE <- rep('SEE', each = 3832)
SR <- rep('SR', each = 3949)
STC <- rep('STC', each = 3992)
STE <- rep('STE', each = 4079)
SEC <- rep('SEC', each = 4051)

strains <- c(SEE, SR, STC, STE, SEC)

total_pca_dataframe_ORF$Strain <- strains
```

Create the PCA, excluding the columns with the extra information 
```{r}
final_orf_PCA <- total_pca_dataframe_ORF[, -c(65:68)] %>% PCA(graph = FALSE)
```
Summary statistics for the PCA 
```{r}
fviz_eig(final_orf_PCA)
get_eigenvalue(final_orf_PCA)
get_pca(final_orf_PCA)
```
black dots pca by orf
```{r}
pdf("ORF_PCA.pdf", width = 9, height = 6)

fviz_pca_ind(final_orf_PCA, geom = "point", title = 'All Open Reading Frames, Components 1 and 2')

final_orf_PCA %>% fviz_pca_ind(c(1,3),geom = "point",  title = 'All Open Reading Frames, Components 1 and 3')

final_orf_PCA %>% fviz_pca_ind(c(2,3),geom = "point",  title = 'All Open Reading Frames, Components 2 and 3')

final_orf_PCA %>% fviz_screeplot()
  
dev.off()
```
PCA for Salmonella serovar Enteritidis, clinical isolate
```{r}
pdf("ORF_SEC_PCA.pdf", width = 9, height = 6)

SEC_pca <- total_pca_dataframe_ORF[c(15854:19903), -c(65:68)] %>% PCA(graph = FALSE)

fviz_pca_ind(SEC_pca,
             geom.ind = "point",
             col.ind = '#E74C3C',
             pointshape = 19,
             title = 'All Open Reading Frames for SEC, Components 1 and 2'
             )
dev.off()
```

PCA for Salmonella serovar Enteritidis, environmental isolate
```{r}
pdf("ORF_SEE_PCA.pdf", width = 9, height = 6)

SEE_pca <- total_pca_dataframe_ORF[c(1:3832), -c(65:68)] %>% PCA(graph = FALSE)

fviz_pca_ind(SEE_pca,
             geom.ind = "point",
             col.ind = '#F39C12',
             pointshape = 19,
             title = 'All Open Reading Frames for SEE, Components 1 and 2'
             )
dev.off()
```

PCA for Salmonella Reference genome
```{r}
pdf("ORF_SR_PCA.pdf", width = 9, height = 6)

SR_pca <- total_pca_dataframe_ORF[c(3833:7781), -c(65:68)] %>% PCA(graph = FALSE)

fviz_pca_ind(SR_pca,
             geom.ind = "point",
             col.ind = '#3498DB',
             pointshape = 19,
             title = 'All Open Reading Frames for SR , Components 1 and 2'
             )
dev.off()
```

PCA for Salmonella serovar Typhimurium, clinical isolate
```{r}
pdf("ORF_STC_PCA.pdf", width = 9, height = 6)

STC_pca <- total_pca_dataframe_ORF[c(7782:11774), -c(65:68)] %>% PCA(graph = FALSE)

fviz_pca_ind(STC_pca,
             geom.ind = "point",
             col.ind = '#0DB615',
             pointshape = 19,
             title = 'All Open Reading Frames for STC, Components 1 and 2'
             )
dev.off()
```

PCA for Salmonella serovar Typhimurium, environmental isolate
```{r}
pdf("ORF_STE_PCA.pdf", width = 9, height = 6)

STE_pca <- total_pca_dataframe_ORF[c(11775:15854), -c(65:68)] %>% PCA(graph = FALSE)

fviz_pca_ind(STE_pca,
             geom.ind = "point",
             col.ind = '#8E44AD',
             pointshape = 19,
             title = 'All Open Reading Frames for STE, Components 1 and 2'
             )
dev.off()
```

Color code all open reading frames by strain
```{r}
pdf("ORF_by_Strain_PCA.pdf", width = 9, height = 6)

total_pca_dataframe_ORF$Strain <- unlist(total_pca_dataframe_ORF$Strain)

fviz_pca_ind(final_orf_PCA,
             geom.ind = "point",
             col.ind = total_pca_dataframe_ORF$Strain,
             palette = c('#E74C3C', '#F39C12', '#3498DB', '#2ECC71', '#8E44AD'),
             legend.title = 'Strains',
             pointshape = 19,
             title = 'All Open Reading Frames by Strain, Components 1 and 2'
             )
final_orf_PCA %>% fviz_pca_ind(c(1,3),
             geom.ind = "point",
             col.ind = total_pca_dataframe_ORF$Strain,
             palette = c('#E74C3C', '#F39C12', '#3498DB', '#2ECC71', '#8E44AD'),
             legend.title = 'Strains',
             pointshape = 19,
             title = 'All Open Reading Frames by Strain, Components 1 and 3'
             )

final_orf_PCA %>% fviz_pca_ind(c(2,3),
             geom.ind = "point",
             col.ind = total_pca_dataframe_ORF$Strain,
             palette = c('#E74C3C', '#F39C12', '#3498DB', '#2ECC71', '#8E44AD'),
             legend.title = 'Strains',
             pointshape = 19,
             title = 'All Open Reading Frames by Strain, Components 2 and 3'
             )
final_orf_PCA %>% fviz_screeplot()

dev.off()
```

Color code all open reading frames by location
```{r}
pdf("ORF_by_Location_PCA.pdf", width = 9, height = 6)

fviz_pca_ind(final_orf_PCA,
             geom = "point",
             col.ind = total_pca_dataframe_ORF$Location,
             gradient.cols = c("#8B66FE", "#669BFE", "#5BFFC8", "#66FE8B", "#5BFFC8", "#669BFE", "#8B66FE"),
             legend.title = 'Location', 
             title = 'Location on Genome, Components 1 and 2'
             )
final_orf_PCA %>% fviz_pca_ind(c(1,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$Location,
             gradient.cols = c("#8B66FE", "#669BFE", "#5BFFC8", "#66FE8B", "#5BFFC8", "#669BFE", "#8B66FE"),
             legend.title = 'Location',
             title = 'Location on Genome, Components 1 and 3'
             )

final_orf_PCA %>% fviz_pca_ind(c(2,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$Location,
             gradient.cols = c("#8B66FE", "#669BFE", "#5BFFC8", "#66FE8B", "#5BFFC8", "#669BFE", "#8B66FE"),
             legend.title = 'Location',
             title = 'Location on Genome, Components 2 and 3'
             )

final_orf_PCA %>% fviz_screeplot()

dev.off()
```

Color code all open reading frames for whether they are in a pathogenicity island
```{r}
pdf("ORF_by_SPI_PCA.pdf", width = 9, height = 6)

fviz_pca_ind(final_orf_PCA,
             geom = "point",
             col.ind = total_pca_dataframe_ORF$SPI,
             palette = c("#D5D4D4", "#FF0000", '#023BFE'),
             legend.title = 'Pathogenicity Island',
             pointshape = 19,
             alpha.ind = 0.5,
             title = 'Pathogenicity Islands One and Two, Components 1 and 2'
             )

final_orf_PCA %>% fviz_pca_ind(c(1,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$SPI,
             palette = c("#D5D4D4", "#FF0000", '#023BFE'),
             legend.title = 'Pathogenicity Island',
             pointshape = 19,
             alpha.ind = 0.5,
             title = 'Pathogenicity Islands One and Two, Components 1 and 3')

final_orf_PCA %>% fviz_pca_ind(c(2,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$SPI,
             palette = c("#D5D4D4", "#FF0000", '#023BFE'),
             legend.title = 'Pathogenicity Island',
             pointshape = 19,
             alpha.ind = 0.5,
             title = 'Pathogenicity Islands One and Two, Components 2 and 3')

final_orf_PCA %>% fviz_screeplot()

dev.off()
```

Color code all open reading frames by GC content
```{r}
pdf("ORF_GC_Content_PCA.pdf", width = 9, height = 6)

fviz_pca_ind(final_orf_PCA,
             geom = "point",
             col.ind = total_pca_dataframe_ORF$GC_content,
             gradient.cols = c("#8695FF", "#E0B3FF", "#FD5D5D"),
             legend.title = 'GC Content',
             title = 'GC Content, Component 1 and 2' 
             )
final_orf_PCA %>% fviz_pca_ind(c(1,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$GC_content,
             gradient.cols = c("#8695FF", "#E0B3FF", "#FD5D5D"),
             legend.title = 'GC Content',
             title = 'GC Content, Component 1 and 3'
             )

final_orf_PCA %>% fviz_pca_ind(c(2,3),geom = "point",
             col.ind = total_pca_dataframe_ORF$GC_content,
             gradient.cols = c("#8695FF", "#E0B3FF", "#FD5D5D"),
             legend.title = 'GC Content',
             title = 'GC Content, Component 2 and 3'
             )

final_orf_PCA %>% fviz_screeplot()

dev.off()
```









